Effects of epidemiological structure on the transient evolution of HIV virulence

### 2017-03-18 ms/ms.tex
dinking around with figures
figure orders and acknowledgement

@@

\begin{figure}[t!p]

\begin{subfigure}[b]{0.4\linewidth}
	\centering
	\raisebox{0.44\height}{
	  \includegraphics[width=0.9\linewidth]{../figures/tikz-f1}
	}
\caption{Instantaneous switching without extra-pair contact (``instswitch'')}
\end{subfigure}
\begin{subfigure}[b]{0.4\linewidth}
	\centering
	  \includegraphics[width=0.9\linewidth]{../figures/tikz-f2}
\caption{Instantaneous switching with extra-pair contact (``instswitch+epc'')}
\end{subfigure}

\begin{subfigure}[b]{0.4\linewidth}
	\centering
	\raisebox{0.33\height}{
	  \includegraphics[width=0.9\linewidth]{../figures/tikz-f3}
	}
\caption{Pair-formation model without extra-pair/uncoupled contact 
(``pairform'')}
\end{subfigure}
\begin{subfigure}[b]{0.4\linewidth}
	\centering
	  \includegraphics[width=0.9\linewidth]{../figures/tikz-f4}
\caption{Pair-formation model with extra-pair/uncoupled contact
(``pairform+epc'')}
\end{subfigure}

\caption{\textbf{Model contact structures.} 
Schematic representations of models with
explicit contact structure. Top row (a,b) shows models with instantaneous
partnership formation; bottom row (c,d)  shows those where individuals remain
single for some period of time between partnerships. Left column (a,c) 
shows models with within-partnership transmission only; in these models
transmission can only occur within serodiscordant partnerships, and
serodiscordant partnerships can only be formed by dissolution
and reformation of partnerships. Right column (b,d) shows models
that allow extra-pair and uncoupled transmission.
\emph{Solid arrows} represent infection transitions; \emph{dotted arrows}
represent partnership dissolution and formation; \emph{dashed arrows}
show influences on infection rate, including both within-pair
transmission (blue) and extra-pair/uncoupled transmission (red).
In addition to these models, we also use random-mixing and implicit
models with only $S$ and $I$ compartments and heterogeneous models
where each compartment is subdivided by sexual activity rate
(of both partners, in the case of partnership compartments).
}
\label{fig:diagrams}
\end{figure}

@@

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig1.pdf}
\caption{\textbf{Baseline dynamics.}
Time series of mean population \Lspvl\ using baseline parameters.
(a) Contrast between the three-stage Shirreff model and the single-stage model calibrated to varying initial exponential growth rates, $r$.
(b) Effects of varying initial infectious density $I(0)$.
(c) Effects of varying initial mean virulence $\alpha(0)$.
The middle $r=0.042$ (orange, dotted) curve in panel (a), calibrated to match the epidemic dynamics of Shirreff \etal's three-stage model \cite{shirreff_transmission_2011}, shows that our simplified single-stage HIV model can produce similar SPVL trajectories to the original three-stage model; the other $r$ values are chosen to show the effects of doubling or halving the initial epidemic growth rate. The other panels (b,c) show that eco-evolutionary dynamics are qualitatively similar across a range of initial conditions.}
\label{fig:panel3}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig2.pdf}
\caption{\textbf{Envelopes of virulence trajectories (population mean \Lspvl) under all models.}
Panels are arranged in order of complexity of contact structure, from 
``random'' (least complex) to ``hetero'' (most complex).
All models were run until $t=4000~\textrm{years}$ using parameters from 1000 Latin hypercube samples to illustrate the range of possible dynamics. Envelopes contain the middle 95\% of trajectories (i.e. we select all points between the 0.025 and 0.975 quantiles for each model at each year), while center lines show mean trajectories. Truncated series (up to year 800) are shown here.}
\label{fig:virtraj}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig3.pdf}
\caption{{\bf Univariate distributions of summary statistics.}
Models arranged in order of complexity, as in \figurename~\ref{fig:virtraj}.
For every simulation, three summary statistics are shown: peak time, maximum mean \Lspvl, and equilibrium mean \Lspvl.
Peak time is the time at which the SPVL reaches its maximum for each simulation run (maximum mean \Lspvl).
Equilibrium mean \Lspvl\ is taken by taking the mean \Lspvl\ value at $t = 4000$.
Because the distribution of equilibrium mean \Lspvl\
(lower left panel) for the random-mixing model is very narrow, it
has been replaced by a point in order to preserve the vertical axis scaling.}
\label{fig:unidist}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig4.pdf}
\caption{{\bf Pairs plot: bivariate relationships among summary statistics for each model structure.}
The three summary statistics for each simulation are plotted against each other in order to visualize the relationship among the summary statistics and to help compare models.
Surprisingly, the implicit model, an approximation for instantaneous partnership formation model (instswitch), shows an almost identical trend with a model that has pair formation dynamics (pairform).
The \emph{dashed line} in the equilibrium vs. peak virulence plot (lower left) shows the 1:1 line, where equilibrium and peak virulence are equal.
To avoid too much overplotting, only 10\% of the parameter sets
(randomly sampled) are shown here.}
\label{fig:pairplot}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig5.pdf}
\caption{{\bf Sensitivity plot.}
Distribution (points) and trend (smooth line) of summary statistics (y-axis,
rows) as a function of parameter values (x-axis, columns), some of which have been scaled to calibrate the initial epidemic growth rate.
Changing parameters may have differing effects for models with different
contact structures: for example, partnership dissolution rate ($c$, third column) increases peak time and virulence in models without explicit pair formation but decreases them in more realistic models.
}
\label{fig:plot_sens}
\end{figure}

@@

@@

\section*{Acknowledgements}
We would like to thank Christophe Fraser and
David Champredon for access to simulation code
and David Earn, Wim Delva, and an anonymous reviewer for useful comments;
this work was funded by NSERC Discovery Grant 386590-2010.
Computational resources were provided by SHARCnet.

>>>>

\begin{figure}[!ht]
\caption{\textbf{Baseline dynamics.}
Time series of mean population \Lspvl\ using baseline parameters.
(a) Contrast between the three-stage Shirreff model and the single-stage model calibrated to varying initial exponential growth rates, $r$.
(b) Effects of varying initial infectious density $I(0)$.
(c) Effects of varying initial mean virulence $\alpha(0)$.
The middle $r=0.042$ (orange, dotted) curve in panel (a), calibrated to match the epidemic dynamics of Shirreff \etal's three-stage model \cite{shirreff_transmission_2011}, shows that our simplified single-stage HIV model can produce similar SPVL trajectories to the original three-stage model; the other $r$ values are chosen to show the effects of doubling or halving the initial epidemic growth rate. The other panels (b,c) show that eco-evolutionary dynamics are qualitatively similar across a range of initial conditions.}
\label{fig:panel3}
\end{figure}

@@

@@

\begin{figure}[t!p]
\caption{\textbf{Model contact structures.} 
Schematic representations of models with
explicit contact structure. Top row (a,b) shows models with instantaneous
partnership formation; bottom row (c,d)  shows those where individuals remain
single for some period of time between partnerships. Left column (a,c) 
shows models with within-partnership transmission only; in these models
transmission can only occur within serodiscordant partnerships, and
serodiscordant partnerships can only be formed by dissolution
and reformation of partnerships. Right column (b,d) shows models
that allow extra-pair and uncoupled transmission.
\emph{Solid arrows} represent infection transitions; \emph{dotted arrows}
represent partnership dissolution and formation; \emph{dashed arrows}
show influences on infection rate, including both within-pair
transmission (blue) and extra-pair/uncoupled transmission (red).
In addition to these models, we also use random-mixing and implicit
models with only $S$ and $I$ compartments and heterogeneous models
where each compartment is subdivided by sexual activity rate
(of both partners, in the case of partnership compartments).
}
\label{fig:diagrams}
\end{figure}

@@

@@

\begin{figure}[!ht]
\caption{\textbf{Envelopes of virulence trajectories (population mean \Lspvl) under all models.}
Panels are arranged in order of complexity of contact structure, from 
``random'' (least complex) to ``hetero'' (most complex).
All models were run until $t=4000~\textrm{years}$ using parameters from 1000 Latin hypercube samples to illustrate the range of possible dynamics. Envelopes contain the middle 95\% of trajectories (i.e. we select all points between the 0.025 and 0.975 quantiles for each model at each year), while center lines show mean trajectories. Truncated series (up to year 800) are shown here.}
\label{fig:virtraj}
\end{figure}

@@

\begin{figure}[!ht]
\caption{{\bf Univariate distributions of summary statistics.}
Models arranged in order of complexity, as in \figurename~\ref{fig:virtraj}.
For every simulation, three summary statistics are shown: peak time, maximum mean \Lspvl, and equilibrium mean \Lspvl.
Peak time is the time at which the SPVL reaches its maximum for each simulation run (maximum mean \Lspvl).
Equilibrium mean \Lspvl\ is taken by taking the mean \Lspvl\ value at $t = 4000$.
Because the distribution of equilibrium mean \Lspvl\
(lower left panel) for the random-mixing model is very narrow, it
has been replaced by a point in order to preserve the vertical axis scaling.}
\label{fig:unidist}
\end{figure}

\begin{figure}[!ht]
\caption{{\bf Pairs plot: bivariate relationships among summary statistics for each model structure.}
The three summary statistics for each simulation are plotted against each other in order to visualize the relationship among the summary statistics and to help compare models.
Surprisingly, the implicit model, an approximation for instantaneous partnership formation model (instswitch), shows an almost identical trend with a model that has pair formation dynamics (pairform).
The \emph{dashed line} in the equilibrium vs. peak virulence plot (lower left) shows the 1:1 line, where equilibrium and peak virulence are equal.
To avoid too much overplotting, only 10\% of the parameter sets
(randomly sampled) are shown here.}
\label{fig:pairplot}
\end{figure}

@@

@@

\begin{figure}[!ht]
\caption{{\bf Sensitivity plot.}
Distribution (points) and trend (smooth line) of summary statistics (y-axis,
rows) as a function of parameter values (x-axis, columns), some of which have been scaled to calibrate the initial epidemic growth rate.
Changing parameters may have differing effects for models with different
contact structures: for example, partnership dissolution rate ($c$, third column) increases peak time and virulence in models without explicit pair formation but decreases them in more realistic models.
}
\label{fig:plot_sens}
\end{figure}

@@

\section*{Acknowledgements}
We would like to thank Christophe Fraser and
David Champredon for access to simulation code
and David Earn for useful comments.
Computational resources were provided by SHARCnet.
